Global patterns of forest autotrophic carbon fluxes

### 2020-03-01 manuscript/manuscript draft.Rmd
worked on results

@@

$R^2$ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2). Of the major fluxes, R~auto~ and GPP were the most strongly explained by latitude and climate, with climate explaining at most 71% of variation in GPP, and 65% in R~auto~. The proportion of variation explained by climate and latitude decreased in NPP and ANPP, with climate explaining at most 51% of variation in NPP and 44% in ANPP. Of the major fluxes, BNPP~root~ was the least well explained by climate and latitude, with climate explaining at most 36% of variation.

With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower. Climate explained at most 24% of the variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~. In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation. This pattern was also seen in the $R^2$ values for multivariate models.

*How do FACF relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on FACF has focused on these as key climate variables. MAT was a significant (p<0.05) and strong ($R^2 \ge$ **##**) predictor of all FACF tested. All FACF increased linearly with temperature (Fig. 4; all p<0.05, $R^2 \ge$ **##**), and we found no support for a saturation point of FACF with temperature.

MAP was found to be a significant (p<0.05) but poor predictor of FACF, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux. For the majority FACF, a polynomial model was the best fit. FACF increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which FACF started to decrease (Fig. 4). The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33. Data above this point was not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.

There was a significant interactive effect between MAT and MAP for NPP and ANPP~woody~ ~stem~ (fig.3). There was a significant additive effect for GPP, ANPP and autotrophic respiration. ANPP~foliage~, BNPP, BNPP~fine~ ~root~, and root respiration showed no significant interactive or additive effect (see table S3).

@@

We found a significant relationship between C flux and potential evapotranspiration for all FACF. ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 4). We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that FACF begin to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of C flux for all FACF. BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig. 4). FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa. At values above 0.8 kPa, FACF began to decrease.

All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation. Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of R~auto~ ($R^2$ = 0.26, p<0.05).

@@

*What is the role of seasonality in explaining FACF?* Temperature seasonality was a significant predictor of FACF. GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (Fig. 4) {**give statistic on signficiance**}. ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality. Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}. FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. In contrast, there was no significant effect of precipitation seasonality on FACF.

>>>>

*How do FACF relate to MAT and MAP?*

All FACF increased linearly with MAT, and we found no support for a saturation point of FACF with MAT (all p<0.05; Figs. 3-4, S4-S5, Table S2). As with latitude, MAT tended to explain more variation in the larger FACF ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_foliage$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{woody-stem}$, $R_{root}$, $BNPP_{root-fine}$; all $R^2$< 0.25).

MAP was a significant (p<0.05) predictor of all FACF but $ANPP_{woody-stem}$ (Figs. 4a, S4-S5; Table S2). However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 37% of variation in FACF. For the majority of FACF, a polynomial model was the best fit. FACF generally increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which they started to decrease (Figs. 4, S4-S5). The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33. Data above this point were not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.

@@

There was a significant interactive effect between MAT and MAP for NPP and $ANPP_{woody-stem}$ (fig.3). There was a significant additive effect for GPP, ANPP and autotrophic respiration. $ANPP_{foliage}$, BNPP, $BNPP_{root-fine}$, and root respiration showed no significant interactive or additive effect (see table S3).

@@

We found a significant relationship between C flux and potential evapotranspiration for all FACF. $ANPP_{foliage}$, $BNPP_{root-fine}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 4). We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that FACF begin to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of C flux for all FACF. $BNPP_{root-fine}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig. 4). FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa. At values above 0.8 kPa, FACF began to decrease.

All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation. Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of$R_{auto}$ ($R^2$ = 0.26, p<0.05).

@@

*What is the role of seasonality in explaining FACF?* Temperature seasonality was a significant predictor of FACF. GPP, NPP, ANPP, and $R_{root}$ exhibited a polynomial relationship with seasonality (Fig. 4) {**give statistic on signficiance**}. $ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality. Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}. FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. In contrast, there was no significant effect of precipitation seasonality on FACF.
